---
title    : "Analyses for 2^nd^ draft"
subtitle : "Cape vs SWA"
author   : "Ruan van Mazijk"
date     : "`r Sys.Date()`"
output   : pdf_document
geometry : a4paper, margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo     = FALSE,
	fig.path = "figures/",
	message  = FALSE,
	warning  = FALSE
)
library(here)
source(here("draft-02/R/setup.R"))
```

# Preamble/outline

Here I layout the "new", second incarnation of the analyses as discussed over the course of May/June 2019, following the first draft of the manuscript.

To reiterate that manuscript, we hypothesise that the greater vascular plant species richness of the GCFR compared to that of the SWAFR is explained by the regions' difference in environmental heterogeneity.

The proposed "story" of questions for the analyses is as follows:

1. Is the GCFR more heterogeneous environmentally than the SWAFR, and does the scale of that heterogeneity differ to that of the SWAFR?
2. Do the regions differ w.r.t. the species richness of both HDS and QDS cells, and, for HDS cells' richness ($S_{HDS}$), does the explanatory power of mean QDS richness ($S_{QDS}$) and turnover ($T_{QDS}$) differ between the regions?
3. Does heterogeneity explain differences in richness and turnover between the regions?

```{r roughness-analyses, child="01_roughness-analyses.Rmd"}
```

```{r turnover-analyses, child="02_turnover-analyses.Rmd"}
```

```{r model-analyses, child="03_model-analyses.Rmd"}
```

```{r appendix, child="04_appendix.Rmd"}
```
