# 2. Species richness & turnover

Do the regions differ w.r.t. the species richness of both HDS and QDS cells, and, for HDS cells' richness ($S_{HDS}$), does the explanatory power of mean QDS richness ($S_{QDS}$) and turnover ($T_{QDS}$) differ between the regions?

To tackle this question, I compare measures of species richness and turnover between the regions. Species richness at the HDS-scale ($S_{\mathrm{HDS}}$) can be partitioned into the average richness of the constituent QDS in HDS ($\overline{S}_{\mathrm{QDS}}$) and species turnover ($T_{\mathrm{QDS}}$) defined^[following Whittaker's original additive definition: $\gamma = \alpha + \beta$] as:

$$
  T_{\mathrm{QDS}} = S_{\mathrm{HDS}} - \overline{S}_{\mathrm{QDS}}
$$

The distributions of these data are presented in Figure 3. To test for significant differences between GCFR and SWAFR values, I use Mann-Whitney $U$-tests and $CLES$ (Table 3), as most of the variables deviate significantly from normality (Shapiro-Wilk normality test; $P < 0.05$).

Additionally, a visualisation of how $S_{\mathrm{HDS}}$ is partitioned into $\overline{S}_{\mathrm{QDS}}$ and $T_{\mathrm{QDS}}$ is presented in Figure 4.

We can conclude that broad scale species richness (i.e. that at the HDS scale) is more strongly driven by turnover between areas (i.e. QDS) than so in the SWAFR.

```{r import-cellular-data, include=FALSE}
HDS <- read_csv(here("draft-02/outputs/QDS_data_cells.csv"))
QDS <- read_csv(here("draft-02/outputs/EDS_data_cells.csv"))

# Remove cells w/ < 4 sub-cells
HDS %<>% filter(n_QDS == 4)
QDS %<>% filter(n_EDS == 4)

# Combine datasets for figure
HDS_QDS <- as_tibble(rbind(
  HDS %>%
    dplyr::select(region, HDS_richness, add_turnover_prop) %>%
    gather(metric, value, HDS_richness, add_turnover_prop),
  QDS %>%
    dplyr::select(region, QDS_richness) %>%
    gather(metric, value, QDS_richness)
))
```

```{r test-richness-turnover-univariate}
HDS_QDS %>%
  mutate(metric = case_when(
    metric == "HDS_richness"      ~ "$S_{\\mathrm{HDS}}$",
    metric == "QDS_richness"      ~ "$S_{\\mathrm{QDS}}$",
    metric == "add_turnover_prop" ~ "$T_{\\mathrm{QDS}} / S_{\\mathrm{HDS}}$"
  )) %>%
  mutate(metric = factor(metric, levels = c(
    "$S_{\\mathrm{HDS}}$",
    "$S_{\\mathrm{QDS}}$",
    "$T_{\\mathrm{QDS}} / S_{\\mathrm{HDS}}$"
  ))) %>%
  group_by(metric) %>%
  summarise(
    P = wilcox.test(value[region == "SWAFR"], value[region == "GCFR"])$p.value,
    CLES_value = CLES(value[region == "SWAFR"], value[region == "GCFR"])
  ) %>%
  mutate_if(is.numeric, ~ifelse(.x < 0.001,
    "$< 0.001$",
    format(round(.x, digits = 3), nsmall = 3)
  )) %>%
  dplyr::select(metric, CLES_value, P) %>%
  knitr::kable(
    caption = paste(
      "Results of Mann-Whitney $U$-tests and the $CLES$ of GCFR vs SWAFR",
      "for various species richness and turnover metrics."
    ),
    col.names = c("Metric", "$CLES$", "$P_U$"),
    align = "lrr"
  )
```

```{r test-richness-turnover-normality, include=FALSE}
# Also show how I tested for metrics deviating from normality
HDS_QDS %>%
  group_by(metric) %>%
  summarise(
    P_GCFR  = shapiro.test(value[region == "GCFR"])$p.value,
    P_SWAFR = shapiro.test(value[region == "SWAFR"])$p.value
  ) %>%
  mutate_at(c("P_GCFR", "P_SWAFR"), list(sig = ~ . < 0.05))
```

```{r visualise-partitions, fig.width=9, fig.height=3, fig.cap="(a) Scatter plot of mean QDS-scale richness ($\\overline{S}_{\\mathrm{QDS}}$) and turnover ($T_{\\mathrm{QDS}}$) with contour lines denoting the $S_{\\mathrm{HDS}}$ that would arise as their sum (i.e. increasing from lower-left to upper-right). Distributions of (a) HDS-scale species richness ($S_{\\mathrm{HDS}}$) and (b) the turnover partition of that richness ($T_{\\mathrm{QDS}} / S_{\\mathrm{HDS}}$)."}
hist_plots <- HDS_QDS %>%
  filter(metric %in% c("HDS_richness", "add_turnover_prop")) %>%
  mutate(metric = case_when(
    metric == "HDS_richness"      ~ "italic(S)[HDS]",
    metric == "add_turnover_prop" ~ "italic(T)[QDS]/italic(S)[HDS]"
  )) %>%
  mutate(metric = factor(metric, levels = c(
    "italic(S)[HDS]",
    "italic(T)[QDS]/italic(S)[HDS]"
  ))) %>%
  ggplot(aes(value, fill = region)) +
    geom_histogram(bins = 20, position = "dodge", colour = "black") +
    scale_fill_manual(name = "Region", values = c("black", "white")) +
    labs(
      # Cheat-x-axis-labels
      x = bquote(
        italic("S")["HDS"]
        ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  # 50 spaces
        italic("T")["QDS"]/italic("S")["HDS"]
      ),
      y = "No. HDS"
    ) +
    facet_grid(~metric, scales = "free_x", labeller = label_parsed) +
    theme(axis.text.y = element_text(angle = 90), strip.text = element_blank())

plot_edges <- HDS %$%
  ceiling(max(c(mean_QDS_richness, add_turnover))) + 10

S_HDS_background <- plot_edges %>%
  {seq(from = 0, to = ., by = 10)} %>%
  {expand.grid(x = ., y = .)} %>%
  mutate(z = x + y)
S_HDS_background_plot <- ggplot(S_HDS_background) +
  lims(x = c(0, plot_edges), y = c(0, plot_edges)) +
  geom_contour(
    mapping     = aes(x, y, z = z),
    binwidth    = 500,
    colour      = "grey90",
    show.legend = TRUE
  ) +
  geom_abline(
    intercept = 0, slope = 1,
    linetype = "dashed", colour = "grey25"
  ) +
  labs(
    x = bquote(italic("T")["QDS"]),
    y = bquote(bar(italic("S"))["QDS"])
  ) +
  theme(
    legend.position = "none",
    axis.text.y     = element_text(angle = 90)
  )

partition_plot <- S_HDS_background_plot +
  geom_point(
    data    = HDS,
    mapping = aes(add_turnover, mean_QDS_richness, fill = region),
    shape   = 21
  ) +
  scale_fill_manual(name = "Region", values = c("black", "white"))

# Add labels to S_HDS contours manually
partition_plot <- partition_plot + geom_text(
  data = tibble(
    add_turnover      = c( 10,   10,   10,  360,  860, 1360),
    mean_QDS_richness = c(500, 1000, 1500, 1650, 1650, 1650),
    HDS_richness      = c(500, 1000, 1500, 2000, 2500, 3000)
  ),
  mapping = aes(add_turnover, mean_QDS_richness, label = HDS_richness),
  angle = -45, vjust = -0.5, colour = "grey50", size = 2.5
)

# Plot panels
plot_grid(
  # Flip partition plot to get axes to line up across panels (b/o text heights)
  partition_plot + coord_flip(),
  hist_plots,
  nrow = 1,
  rel_widths = c(1, 2)
)
```
