---
title: "Artificially high turnover between low richness communities"
author: "Ruan van Mazijk"
date: "13 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, 
    message = FALSE, 
    warning = FALSE, 
    results = "hide"
)
```

```{r}
source("turnover_across_richness.R")
```

# Introduction

Here I explore the surfaces of turnover (= Jaccard's distance) values between two communities when varying the richness of those two communities. These simulated simple communities are each drawn from regional species pools (with a specificed richness) without replacement to a predefined richness.

# Max community richness $\approx$ regional richness

```{r}
turnover_across_richness(100, 100)
```

```{r}
turnover_across_richness(500, 500)
```

# Max community richness $\lt$ regional richness

## Regional richness = 1000

```{r}
turnover_across_richness(750, 1000)
```

```{r}
turnover_across_richness(500, 1000)
```

```{r}
turnover_across_richness(250, 1000)
```

```{r}
turnover_across_richness(100, 1000)
```

## Regional richness = 500

```{r}
turnover_across_richness(500, 500)
```

```{r}
turnover_across_richness(250, 500)
```

```{r}
turnover_across_richness(50, 500)
```


## Regional richness = 100

```{r}
turnover_across_richness(75, 100)
```

```{r}
turnover_across_richness(50, 100)
```

```{r}
turnover_across_richness(25, 100)
```

```{r}
turnover_across_richness(10, 100)
```

<!-- TODO: re knit, and also run last chunk in loop to replicate its fuzzy pattern and unfuzzy it??-->
